# Week 3
## Lesson monday 2020-11-09
generalized linear models: works also for example for binary response variables

```{r}
library(tidyverse)
alc <- read.csv("data/alc.csv")
colnames(alc)
```

Quick way to plot counts of a lot of variables
```{r}
gather(alc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```

## Datacamp video; logistic regression
## Datacamp video; odds ratio
odds ratio = odds of Y=1 when has property X / odds of Y=1 when has NO X

"study the relationships between high/low alcohol consumption and some of the other variables in the data. To do this, choose 4 interesting variables in the data and for each of them, present your personal hypothesis about their relationships with alcohol consumption. "
1) freetime: more free time leads to higher alcohol use
2) goout: going out leads to higher alcohol use
3) address: students with rural home environments drink less
4) Medu: students with highly educated mothers drink less

```{r}
g1 <- ggplot(alc, aes(x = high_use, y = freetime)) 
g1 + geom_boxplot()
g2 <- ggplot(alc, aes(x = high_use, y = goout)) 
g2 + geom_boxplot()
g4 <- ggplot(alc, aes(x = high_use, y = Medu)) 
g4 + geom_boxplot()

with(alc, table(high_use, address))
```

## the model:
```{r}
m <- glm(high_use ~ freetime + goout + Medu + address, data = alc, family = "binomial")
# compute odds ratios (OR)
OR <- coef(m) %>% exp

# compute confidence intervals (CI)
CI <- confint(m) %>% exp

# print out the odds ratios with their confidence intervals
cbind(OR, CI)
```



