complete.cases() returns true if case has no NAs

correlation matrix cor()
correlation plot: cor(df) %>% corplot()

Video: dimensionality reduction techniques
dimensionality ~~ number of variables per observation

PCA
components are uncorrelated to eachother
1st captures most variance, 2nd most of remaining variance etc. -> *Order of importance*
assumes that features(variables) with mre variance are more important -> standardization of data

"There are two functions in the default package distribution of R that can be used to perform PCA: princomp() and prcomp(). The prcomp() function uses the SVD and is the preferred, more numerically accurate method."

standardization of data: scale()

prcomp(df) %>% biplot()

-> scatter plot of observations according to PC1 and PC2
arrows represet the original features 
  -> pointing same way = pos. correlation
    -> 90 deg angle = no correlation
  -> pointing same way as axis: correlated to the PC
  
Multiple correspondence analysis (MCA): like PCA but for categorical variables

```{r}
library(FactoMineR)
data(tea)
# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))

# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)

# summary of the model
summary(mca)

# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
```

Then onto the actual course exercise

```{r echo = TRUE, message = FALSE }
library(tidyverse)
library(GGally)
human <- read_csv("data/human.csv")
ggpairs(human)
```

```{r}
summary(human)
```

```{r}
pca <- prcomp(human)
biplot(pca)

#here we see that the first principal component captures nearly all of the variability of the data. THis is due to the data not being standardizet properly i.e. the variances of the variables being of different sizes. 
round(100 * summary(pca)$importance[2,], digits = 2)
```

Now with scaled (standardized data)
```{r}
humanStd <- scale(human)
pca <- prcomp(humanStd)
biplot(pca)

round(100 * summary(pca)$importance[2,], digits = 2)
```

```{r}
library(FactoMineR)
data(tea)
str(tea)
dim(tea)

keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- select(tea, one_of(keep_columns))
mca <- MCA(tea_time)
plot(mca, invisible = c("ind"))
```

